---
title: "Exam Statistics 2024_3"
author: Your Name
format: docx
editor: visual
markdown: 
  wrap: 60
---

```{r}
#any packages needed? put them here!
```

Please give your answers underneath the question, referencing question numbers.

Please format text answers as comments or put them in text parts.

# Q1: Please generate the following sequences of numbers:

## Q1a: Integers between 95 and 45 in descending order, including borders

```{r}

```

## Q1b: 50 random numbers from a uniform distribution with minimum=18 and maximum=85

```{r}

```

## Q1c: 250 random numbers from a Normal distribution with mean=10 and SD=2

```{r}

```

## Q1d: 20 random numbers from a Poisson distribution with lambda=5

```{r}

```

## Q1e: 6 Lotto numbers (unique, uniform distribution) between 1 and 49 ("6 aus 49").

```{r}

```

## Q2: Test the various numbers you generated in Q1 against the Normal distribution, create a table structure with the p-values and your interpretation

```{r}

```

# Q3: Please explain the following terms in your own words (rather than just copy/pasting from wikipedia/chatbots)

## Q3a: standard deviation

## Q3b: standard error of the mean

## Q3c: significance (as in 'we found a *significant* difference)

## Q3d: statistical power

## Q3e: sensitivity

## Q3f: specificity

# Q4: Please name the class / function for test statistics in the following cases:

## Q4a: Comparison of mean values for 2 groups for a Gaussian measure

## Q4b: Comparison of central tendencies for ordinal data between two groups; what descriptive statistic would be suitable together with the test?

# Q5: Please import and analyze the dataset bordeaux.xlsx

Info on dataset:

-   year: year of harvest

-   temperature: sum of daily average temperatures (in Celsius degrees)

-   sun: duration of insolation (in hours)

-   heat: number of super-hot days

-   rain: rain level (in millimeters)

-   quality: wine quality: a factor with levels bad, good, and medium

## Q5a: Program the import and "pick" / address the first 5 rows from columns rain and quality

```{r}

```

## Q5b: Create a box plot for rain level, grouped by quality, using ggplot. Try to use a reasonable order of quality levels

```{r}

```

## Q5c: Test, if the group differences you observe in Q5b could be a chance outcome, formulate the appropriate Null hypothesis. (keep in mind that '*dependent variable*' does not imply causality)

```{r}

```

## Q5d: Create a report table with descriptive statistics for temperature and rain for the levels of quality.

```{r}

```

## Q5e: Export that table into a text file with a ';' as separator

```{r}

```

# Q6: Create a matrix named 'participants' with 3 columns and 20 rows.

-   Fill column 1 with the words "participant 1" to "participant 20"

-   Put your first name in row 2 column 2 and your family name in row 2 column 3

```{r}

```

# Q7: Program a loop that runs along the rows of matrix from Q6 and that prints the run number and the first name

```{r}

```

# 

***Good luck!***

# 
